Reanalysis Step 4: Propensity Score Analysis with Twang

### 2025-09-03 reanalysis-4-propensity_scores.qmd
deparallelize

library(twang)
library(ggplot2)
library(gridExtra)
library(knitr)
library(parallel)
library(doParallel)

@@

# PARALLEL PROCESSING CONFIGURATION
max_cores_for_ps <- 1  # Maximum cores to use for propensity score estimation (reduced for memory management)

@@

# Set up parallel processing
# Reduce cores for memory-intensive GBM models
available_cores <- detectCores()
log_message(paste0("Detected ", available_cores, " available cores"))
log_message(paste0("Max cores configured: ", max_cores_for_ps))
# Use fewer cores to avoid memory issues - each GBM model needs substantial memory
# Don't subtract 1 from max_cores_for_ps since it's already the desired limit
n_cores <- if (available_cores > max_cores_for_ps) max_cores_for_ps else max(1, available_cores - 1)
log_message(paste0("Parallel processing: Using ", n_cores, " cores (of ", available_cores, " available) to manage memory usage"))

@@

# Create assignments of imputations to workers
# With 2 workers: Worker 1 gets 1-15, Worker 2 gets 16-30
worker_assignments <- data.frame(
  imputation = 1:n_imputations,
  worker_id = rep(1:n_cores, length.out = n_imputations)[1:n_imputations]
)

# For more even distribution with 2 workers
if (n_cores == 2) {
  midpoint <- ceiling(n_imputations / 2)
  worker_assignments$worker_id <- ifelse(worker_assignments$imputation <= midpoint, 1, 2)
}

# Run parallel processing
log_message(paste0("Starting parallel processing of ", n_imputations, " imputations using ", n_cores, " cores"))
log_message(paste0("Worker 1 will process imputations: ", 
                   paste(worker_assignments$imputation[worker_assignments$worker_id == 1], collapse = ", ")))
if (n_cores >= 2) {
  log_message(paste0("Worker 2 will process imputations: ", 
                     paste(worker_assignments$imputation[worker_assignments$worker_id == 2], collapse = ", ")))
}
parallel_start <- Sys.time()

# Use mclapply on Unix/Mac, parLapply on Windows
if (.Platform$OS.type == "unix") {
  # Unix/Mac - use mclapply
  parallel_results <- mclapply(
    1:n_imputations,
    function(i) {
      worker_id <- worker_assignments$worker_id[i]
      process_imputation(i, imputed_datasets, ps_formula, n_trees, cache_dir, use_cache, worker_id, log_dir)
    },
    mc.cores = n_cores
  )
} else {
  # Windows - use parLapply
  log_message(paste0("Creating cluster with ", n_cores, " workers"))
  cl <- makeCluster(n_cores)
  log_message(paste0("Cluster created with ", length(cl), " workers"))

  # Export necessary objects to cluster
  clusterExport(cl, c(
    "process_imputation", "imputed_datasets", "ps_formula", 
    "n_trees", "cache_dir", "use_cache", "worker_assignments", "log_dir"
  ), envir = environment())

  # Load required packages on each worker
  clusterEvalQ(cl, {
    library(twang)
  })

  # Create wrapper function for parLapply
  parallel_results <- parLapply(cl, 1:n_imputations, function(i) {
    worker_id <- worker_assignments$worker_id[i]
    process_imputation(i, imputed_datasets, ps_formula, n_trees, cache_dir, use_cache, worker_id, log_dir)
  })

  stopCluster(cl)
}

parallel_time <- round(as.numeric(difftime(Sys.time(), parallel_start, units = "mins")), 2)
log_message(paste0("Parallel processing completed in ", parallel_time, " minutes"))

@@

for (res in parallel_results) {
  twang_results[[res$imputation]] <- res$result

@@

# Read and consolidate worker logs
log_message("=== Worker Log Summary ===")
for (worker in 1:n_cores) {
  worker_log_file <- file.path(log_dir, paste0("propensity_scores_worker_", worker, ".log"))
  if (file.exists(worker_log_file)) {
    worker_logs <- readLines(worker_log_file)
    log_message(paste0("Worker ", worker, " processed ", 
                      sum(grepl("started", worker_logs)), " imputations, ",
                      sum(grepl("completed", worker_logs)), " completed"))
    # Optionally show first and last entries
    if (length(worker_logs) > 0) {
      log_message(paste0("  First: ", worker_logs[1]))
      log_message(paste0("  Last:  ", worker_logs[length(worker_logs)]))
    }
  }
}
log_message("==========================")

@@

# Display progress summary with caching and timing information
kable(progress_df, 
      caption = paste0("Twang Model Fitting Progress (Total time: ", parallel_time, " minutes using ", n_cores, " cores)"),
      col.names = c("Imputation", "Worker", "Trees Used", "Min Weight", "Max Weight", "From Cache", "Time (min)"))
```

>>>>

library(twang)
library(ggplot2)
library(gridExtra)
library(knitr)

@@

# SEQUENTIAL PROCESSING (parallel clustering removed for stability)

@@

# Sequential processing - no parallel cores used
log_message("Using sequential processing (parallel clustering removed for stability)")

@@

# Run sequential processing (parallel clustering removed due to stability issues)
log_message(paste0("Starting sequential processing of ", n_imputations, " imputations"))
sequential_start <- Sys.time()

# Process each imputation sequentially
sequential_results <- lapply(
  1:n_imputations,
  function(i) {
    worker_id <- 1  # Single worker for sequential processing
    process_imputation(i, imputed_datasets, ps_formula, n_trees, cache_dir, use_cache, worker_id, log_dir)
  }
)

sequential_time <- round(as.numeric(difftime(Sys.time(), sequential_start, units = "mins")), 2)
log_message(paste0("Sequential processing completed in ", sequential_time, " minutes"))

@@

for (res in sequential_results) {
  twang_results[[res$imputation]] <- res$result

@@

# Read worker log (single sequential worker)
log_message("=== Processing Log Summary ===")
worker_log_file <- file.path(log_dir, "propensity_scores_worker_1.log")
if (file.exists(worker_log_file)) {
  worker_logs <- readLines(worker_log_file)
  log_message(paste0("Sequential processing: ", 
                    sum(grepl("started", worker_logs)), " imputations started, ",
                    sum(grepl("completed", worker_logs)), " completed"))
  # Optionally show first and last entries
  if (length(worker_logs) > 0) {
    log_message(paste0("  First: ", worker_logs[1]))
    log_message(paste0("  Last:  ", worker_logs[length(worker_logs)]))
  }
}
log_message("==========================")

@@

# Display progress summary with caching and timing information
kable(progress_df, 
      caption = paste0("Twang Model Fitting Progress (Total time: ", sequential_time, " minutes - sequential processing)"),
      col.names = c("Imputation", "Worker", "Trees Used", "Min Weight", "Max Weight", "From Cache", "Time (min)"))
```
